---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Loading and preprocessing the data

```{r}
setwd("~/2.EDUCATION/DataScience/ReproducibleResearch/RepData_PeerAssessment1")
unzip("activity.zip")
library(foreign)
act.data <- read.csv("activity.csv")
```


## What is mean total number of steps taken per day?

```{r}
require(dplyr)
day.steps <- act.data %>% group_by(date) %>% summarise(sum(steps, na.rm = TRUE))
names(day.steps)[2] <- "steps"
mean.day.steps <- mean(day.steps$steps, na.rm = TRUE)
median.day.steps <- median(day.steps$steps, na.rm = TRUE)
```

The average total number of steps taken per day is `r mean.day.steps` steps and the median is `r median.day.steps` steps.

This is the histogram of the total steps taken daily: 

```{r}
library(ggplot2)
g <- ggplot(day.steps, aes(x = steps)) + geom_histogram()
g <- g + geom_vline(aes(xintercept = mean(steps, na.rm = TRUE)), color = "red", linetype = "dashed", size = 1, show.legend = TRUE) + geom_vline(aes(xintercept = median(steps, na.rm = TRUE)), color = "blue", linetype = "dashed", size = .5, show.legend = TRUE) + scale_color_manual(name = "statistics", values = c(median = "blue", mean = "red")) 
g <- g + ggtitle("Number of Steps Taken by Day") + theme(plot.title = element_text(lineheight=.8, face="bold"))
g
```

## What is the average daily activity pattern?
```{r}
interval.steps <- act.data %>% group_by(interval) %>% summarise(mean(steps, na.rm = TRUE))
names(interval.steps)[2] <- "mean.steps"
max <- interval.steps[which.max(interval.steps$mean.steps),]
```


The time series of the 5-min interval and the average number of steps taken looks like this:

```{r}
q <- ggplot(interval.steps, aes(x=interval, y=mean.steps)) + geom_line(colour= "blue")
q
```

An the 5-minute interval that contains the maximum number of steps (on average across all the days in the dataset) is interval `r max$interval` min, with `r max$mean.steps` steps.


## Imputing missing values

The number of rows with missing values is `r sum(is.na(act.data$steps))`. 


Imputing missing values and new dataset: 

```{r}
library(plyr)
act.data.filled <- act.data
act.data.filled$newsteps <- act.data.filled$steps
mergedata <- merge(act.data.filled, interval.steps, by = "interval")
son.na <- is.na(mergedata$newsteps)
mergedata[son.na, "newsteps"] <- mergedata[son.na,"mean.steps"]
head(mergedata)
```

```{r}
detach(package:plyr)    
library(dplyr)
day.steps2 <- mergedata %>% group_by(date) %>% summarise(suma.nueva = sum(newsteps))
```

Histogram:

```{r}
j <- ggplot(day.steps2, aes(suma.nueva)) + geom_histogram(bins = 50)
j <- j + geom_vline(aes(xintercept = mean(suma.nueva)), color = "red", linetype = "dashed", size = 1, show.legend = TRUE) + geom_vline(aes(xintercept = median(suma.nueva)), color = "blue", linetype = "dashed", size = .5, show.legend = TRUE) + scale_color_manual(name = "statistics", values = c(median = "blue", mean = "red")) 
j <- j + labs(x="Daily Steps")
j
```


## Are there differences in activity patterns between weekdays and weekends?

Create new factor, weekday and weekends:

```{r}
mergedata$date <- as.Date(mergedata$date)
mergedata$wd <- ifelse(weekdays(mergedata$date) %in% c("Saturday", "Sunday"), "weekend", "weekday")
week.noweek <- mergedata %>% group_by(wd, interval) %>% summarise(promedio = mean(newsteps))
k <- ggplot(week.noweek, aes(x=interval, y=promedio)) + geom_line(colour="red")
k <- k + facet_grid(wd~.)
k
```

